% Chapter Template

\chapter{Literature Review} % Main chapter title

\label{Chapter2} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

1. look for relevant scholarly resources
2. evaluate the resources 
    What problem is addressed in the resource?
    How has the author defined the main concepts?
    What theories and methods are used in the resource?
    What is the conclusion of the resource?
    What is the relationship between the resource and other resources?
    How does the resource contribute to knowledge about the topic?

3. identify debates and gaps in these resources
    Patterns and trends, especially in theories, methods, and results.
    Debates, major conflicts, and contradictions.
    Themes.
    Gaps on what is missing in the literature.
    Pivotal publications.


4. develop your outline
order: chronological, thematic, methodological, theoretical
    

5. write literature review
    Introduction
        give focus of the literature review
    body
        finer details.
        summerize, analyze, interpret
        evaluate comprehensively
        write carefully in properly structured and easy to read paragraphs
        
    conclusion

%-----------------------------------
%	SUBSECTION 1
%-----------------------------------

\section{Main Section 1}
\subsection{Subsection 1}


%-----------------------------------
%	SUBSECTION 2
%-----------------------------------

\subsection{Subsection 2}

%----------------------------------------------------------------------------------------
%	SECTION 2
%----------------------------------------------------------------------------------------

\section{Main Section 2}
